---
title: "Solution"
subtitle: "Aesthetics - Data Visualization - Part 2"
output: html_notebook
---

### Exercise 1
Besides `color`, one could also map `class` to other aesthetics such as `shape`, `size`, and `alpha`. As a practice, using the code from the notebook try to regenerate the plot experimenting with different mappings. For instance:

1. Use `fl` as "shape" for your scatterplot.
2. Use `cyl` as the "size" of your points.
3. Use `class` as "alpha" (transparency) for the points.

```{r include=FALSE}
library(tidyverse)
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = fl))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cyl))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

----

### Exercise 2
What's gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

----

### Exercise 3
Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

```{r}
glimpse(mpg)
```
Alternatively, if you are using RStudio then by just type mpg the tibble should give you the variable type information.

```{r}
mpg
```


----

### Exercise 4
Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point()
```
Instead of using discrete colors, the continuous variable uses a scale that varies from a light to dark blue color.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, size = cty)) +
  geom_point()
```

```{r}
#ggplot(mpg, aes(x = displ, y = hwy, shape = cty)) +
#  geom_point()
# Error: A continuous variable can not be mapped to shape
```

----

### Exercise 5
Try mapping the same variable to multiple aesthetics; what do you observe? For intance, in the `displ` vs. `hwy` chart use `hwy` for size and color.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = hwy, size = displ)) +
  geom_point()
```

----

### Exercise 6
What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)

Stroke changes the size of the border for shapes (21-25). These are filled shapes in which the color and size of the border can differ from that of the filled interior of the shape.

For example:

```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, color = "black", fill = "white", size = 5, stroke = 5)
```

----

### Exercise 7
What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`?

Aesthetics can also be mapped to expressions (code like `displ < 5`). It will create a temporary variable which takes values from the result of the expression. In this case, it is logical variable which is `TRUE` or `FALSE`. 

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = displ < 5)) +
  geom_point()
```
