---
title: "Exercise"
subtitle: "Aesthetics - Data Visualization - Part 2"
output: html_notebook
---

### Exercise 1
Besides `color`, one could also map `class` to other aesthetics such as `shape`, `size`, and `alpha`. As a practice, using the code from the notebook try to regenerate the plot experimenting with different mappings. For instance:

1. Use `fl` as "shape" for your scatterplot.
2. Use `cyl` as the "size" of your points.
3. Use `class` as "alpha" (transparency) for the points.

```{r}
# Your answer goes here 
```

----

### Exercise 2
What's gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

```{r}
# Your answer goes here
```

----

### Exercise 3
Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

```{r}
# Your answer goes here
```

----

### Exercise 4
Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
# Your answer goes here
```

----

### Exercise 5
Try mapping the same variable to multiple aesthetics; what do you observe? For intance, in the displ vs. hwy chart use hwy for size and color.

```{r}
# Your answer goes here
```

----

### Exercise 6
What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)

```{r}
# Your answer goes here
```

----

### Exercise 7
What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`?

```{r}
# Your answer goes here
```
