---
title: "Exercise"
subtitle: "Data Transformation - Part 1"
output: html_notebook
---

```{r echo=FALSE}
# loading libraries
library(tidyverse)
library(nycflights13)

# modifying chart size
options(repr.plot.width=5, repr.plot.height=3)
```

### Exercise 1
Find all flights that

1. Had an arrival delay of two or more hours
2. Flew to Houston (IAH or HOU)
3. Were operated by United, American, or Delta (hint: use `airline` dataset to get the two-character carrier codes)
4. Departed in summer (July, August, and September)
5. Arrived more than two hours late, but didn't leave late
6. Departed by at least an hour delay, but made up over 30 minutes in flight
7. Departed between midnight and 6am (inclusive)

**Answer:**

---

### Exercise 2
Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous exercise, part 4?

**Answer:**

---

### Exercise 3
How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?

**Answer:**

---

### Exercise 4
1) How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()` and apply your answer to the `df` dataframe defined below):

```{r}
(df <- tibble(x = c(5, 6, 2, NA)))
```

2) Apply this answer to `flights` dataset to list the canceled flights first (assuming `NA` for `dep_delay` means the flight was canceled) and then show other flights starting with the ones with large `dep_delay` in a descending order.

**Answer:**

---

### Exercise 5
Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.

**Answer:**

---

### Exercise 6
What happens if you include the name of a variable multiple times in a `select()` call?

**Answer:**

---

### Exercise 7
Does the result of running the following code surprise you? How do the `select` helpers deal with case by default? Can you change that default?

```{r}
select(flights, contains("TIME"))
```

**Answer:**
