---
title: "Data Transformation"
subtitle: "Part 2"
output: html_notebook
---

### Table of Contents

* Add new variables with `mutate()`
* Grouped summaries with `summarize()`
* Group by with `group_by()`

```{r echo=FALSE}
# loading libraries
library(tidyverse)
library(lubridate)
library(nycflights13)
library(plotly)

# modifying chart size
options(repr.plot.width=5, repr.plot.height=3)
```

## Add new variables with `mutate()`
Besides selecting sets of existing columns, it's often useful to add new columns that are functions of existing columns. That's the job of `mutate()`.

`mutate()` always adds new columns at the end of your dataset, but doesn't change the original dataframe. If you want to keep the outcome of `mutate()`, or any of the other functions we discussed in the previous notebook, assign the output to a new or existing object.

```{r}
# Starting with a smaller number of columns
flights_sml <- select(flights, 
                      year:day, 
                      ends_with("delay"), 
                      distance, 
                      air_time
                     )

flights_sml
```

```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
      )
```

Alternatively we could've used our piping skills:

```{r}
flights %>%
  select(year:day, 
         ends_with("delay"), 
         distance, 
         air_time
        ) %>%
  mutate(gain = dep_delay - arr_delay,
         speed = distance / air_time * 60)
```

Note that you can refer to columns that you've just created:

```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours
      )
```

---

## Grouped summaries with `summarize()`
The last key verb is `summarize()` (or `summarise()`). It collapses a data frame to a single row:

```{r}
# Average delay
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

What happens if we don't specify `na.rm = TRUE`?

### `group_by()`

`summarize()` is not terribly useful unless we pair it with `group_by()`. This changes the unit of analysis from the complete dataset to individual groups. Then, when you use the dplyr verbs on a grouped data frame they'll automatically be applied "by group". For example, if we applied exactly the same code to a data frame grouped by date, we get the average delay per date:

```{r}
# Average delay for each day
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE))
```

### Examples

How does the average distance to destination impact the arrival delay on average?

```{r}
delay <- flights %>% 
  group_by(dest) %>%  # group by destination
  summarize(count = n(),  # counting the number of flights (per destination)
            Distance = mean(distance, na.rm = TRUE),  # average distance (per destination)
            Delay = mean(arr_delay, na.rm = TRUE)  # average delay (per destination)
           ) %>%
  filter(count > 20, dest != "HNL")  # keeping destinations with more than 20 flights that are not "HNL"

delay
```

```{r}
p <- ggplot(data = delay, mapping = aes(x = Distance, y = Delay)) +
  geom_point(aes(size = count, color = dest), alpha = 1/3) +
  geom_smooth(se = FALSE) +
  labs(title = "Average distance to destination vs. arrival delay",
      caption = "The circle size shows the number of flights to that destination.")

ggplotly(p + theme(legend.position = "none"))  # Same plot with ggplotly()
```

It looks like delays increase with distance up to ~750 miles and then decrease. Maybe as flights get longer there's more ability to make up delays in the air?

Can you figure out why we removed "HNL" (Pacific/Honolulu)?

---

Let's examine if the number of flights during the summer has increased:

```{r}
flight_count <- flights %>%
  mutate(date = make_datetime(year, month, day)) %>%
  group_by(date) %>%
  summarize(count = n(),  # counting the number of flights (per day)
            delay = mean(arr_delay, na.rm = TRUE)  # average delay (per day)
           )

p <- ggplot(flight_count, aes(date, count)) +
  geom_line() +
  labs(title = "Number of flights out of NYC") +
  theme_classic()

ggplotly(p)
```

Interestingly, all of these dips are weekends. Let's check this by creating a bar chart:

```{r}
flights %>%
  mutate(date = make_datetime(year, month, day),
        weekday = wday(date, label = TRUE)) %>%
  group_by(weekday) %>%
  summarize(count = n(),  # counting the number of flights (per weekday)
            delay = mean(arr_delay, na.rm = TRUE)  # average delay (per weekday)
           ) %>% 
  ggplot(aes(weekday, count)) + 
    geom_bar(stat="identity", width = 0.5) + 
    labs(title="Number of flights per dat of week",
        x = "Day of Week",
        y = "Count") +
    theme_classic() +
    theme(axis.text.x = element_text(angle=65, vjust=0.6))
```
