---
title: "Solution"
subtitle: "Relational Data - Data Manipulation (Part 2)"
output: html_notebook
---

```{r echo=FALSE}
# loading libraries
library(tidyverse)
library(nycflights13)

# modifying chart size
options(repr.plot.width=5, repr.plot.height=3)
```

### Exercise 1
Identify the keys in ggplot2::diamonds

```{r}
# Your answer goes here
```

### Exercise 2
Add the location of the origin and destination (i.e. the `lat` and `lon`) to flights.

```{r}
# Your answer goes here.
```

---

### Exercise 3
What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

Hint: Use the following code as a template for creating a simple US map and annotate it useful information. 

Note: You may need to install/load **maps** library.

```{r}
airports %>%
  filter(lon < 0 & between(lat, 25, 50)) %>%  # limiting the map to US mainland boarders for clarity
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(size = 0.5) +
    coord_quickmap()
```

```{r}
# Your answer goes here.
```

### Exercise 4
What does it mean for a flight to have a missing `tailnum`? What do the tail numbers that don't have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

```{r}
# Your answer goes here
```

---

### Exercise 5
Filter `flights` to only show flights with planes that have flown at least 100 flights.

```{r}
# Your answer goes here
```

---

### Exercise 6
* What does `anti_join(flights, airports, by = c("dest" = "faa"))` tell you? 
* What does `anti_join(airports, flights, by = c("faa" = "dest"))` tell you?

```{r}
# Your answer goes here
```