---
title: "CSV & Excel Read/Write"
subtitle: "Data Import/Export"
output: html_notebook
---

```{r include=FALSE}
# loading libraries
library(tidyverse)

# modifying chart size
options(repr.plot.width=5, repr.plot.height=3)
```


### Table of Contents
* Reading flat files into data frames
* Reading from inline
* Parsing a vector
* Writing to a file
* Read Excel files
* Write to an Excel file

As you saw in the earlier notebooks some of R packages provide sample data for educational purposes, for instance `mpg` and `diamonds`. In reality we would want to import our own data for analysis. Base R provides functions for reading and exporting data files but it's not as convenient, and fast as some of the other packages available to us. Here we will use the package **readr**, it is a part of *tidyverse* suite, so, no need to install it since we already have tidyverse. In order to use it we need to load tidyverse (or directly `readr`).

## Reading flat files into data frames
Under the *data* folder we have a comma-separated value (CSV) file called `Telco-Customer-Churn.csv`. Let's try to load it using `read_csv()` function from readr. Before doing so, let's check our current working directory by:

```{r}
getwd()
```

Yours will be different in two ways. 1) the path leading to `Intro-to-Data-Analytics-in-R` folder is different. 2) the folder containing your notebook is in the `sandbox` folder. The following code cell will fail, unless you modify yours to reflect the correct relative path:

```
telcoChurnDf <- read_csv("../../data/Telco-Customer-Churn.csv")
```

```{r}
telcoChurnDf <- read_csv("../data/Telco-Customer-Churn.csv")
```

```{r}
telcoChurnDf
```

This behavior is different when you use an R script within a project. There, you can provide an absolute path from your project's root. In that scenario you will provide `"data/Telco-Customer-Churn.csv"` and this will work from anywhere in your project. No more need to modify the path when moving the code or notebook around.

As it appears from the look of printed data frame we immediately realize that `read_csv()` has automatically loaded our CSV file into a tibble data frame and associated it with the name we provided. We can check its class:

```{r}
class(telcoChurnDf)
```

If we look at the help page for `read_csv()` we see what defaults were used for importing the file:

> read_csv(file, col_names = TRUE, col_types = NULL, locale = default_locale(), na = c("", "NA"), quoted_na = TRUE, quote = "\"", comment = "", trim_ws = TRUE, skip = 0, n_max = Inf, guess_max = min(1000, n_max), progress = show_progress())
  
To get details for each of these parameters scroll down to see descriptions and examples.

We started with `read_csv()` since it's by far the most popular *flat file*. But not all the rectangular files have their fields separated by a comma (","). Most common *delimiter-separated values* (DSV) files are: comma (",", also CSV), tab ("    ", also TSV), colon (";"), and pipe ("|"). `read_delim()` is a generic function that can handle different delimiters, note that the second argument `delim` must be provided:


> read_delim(file, delim, quote = "\"", escape_backslash = FALSE, escape_double = TRUE, col_names = TRUE, col_types = NULL, locale = default_locale(), na = c("", "NA"), quoted_na = TRUE, comment = "", trim_ws = FALSE, skip = 0, n_max = Inf,
  guess_max = min(1000, n_max), progress = show_progress())

There are several different functions to support other formats such as fixed width files, and log files. To learn more visit [readr.tidyverse.org](https://readr.tidyverse.org/)

## Reading from inline 
Any of these functions will work with reading inline instead of a file. This is primarily useful for experimentation:

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```

```{r}
sampleDf <- read_csv("a,b,c
1,2,3
4,5,6")
sampleDf
```

```{r}
calDf <- read_delim("First 2 lines to be skipped!
Calendar for September 2019
Mon|Tue|Wed|Thu|Fri|MONTH
2|3|4|5|6|September
9|10|11|12|13|September
16|17|18|19|20|September
23|24|25|26|27|September
30|--|--|--|--|September",
delim = "|",
skip = 2,
na = "--")

calDf
```

The data might not come with column name, we can tell the function that don't use the first row as column name by `col_names = FALSE`, which is the default behavior:

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

`read_csv()` is providing names for the columns, we can also specify them manually:

```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

---

## Writing to a file
Similar to how we read a file from disk we can write one. We can use `write_csv()`, `write_delim()`, and other functions depending on our need. Earlier in this notebook we created `calDf`, a data frame from an inline text which was delimited by "|". Let's write this data frame to file, we can use any of the common delimiters, I pick CSV here:

```{r}
calDf
```

```{r}
write_csv(calDf, path = "calDf.csv")
```

```{r}
# Reading the first 3 rows from disk
calDfNew <- read_csv("calDf.csv", n_max = 3)
calDfNew
```

---

## Read Excel files

We will be using package readxl from tidyverse, but is not a core package so it needs to be loaded explicitly:

```{r}
library(readxl)  # for reading Excel files
```

We can check out the sheet names with `excel_sheets()`:

```{r}
myExcelPath <- "../data/datasets.xlsx"
excel_sheets(myExcelPath)
```

and read any of these sheets into a data frame by `read_excel()`:

```{r}
mtcarsDf <- read_excel(myExcelPath, sheet = 'mtcars')
mtcarsDf
```

if `sheet` parameter is not specified it'll go to the default which is the first sheet.

We can read only a subset of the cells by an Excel-specified notation:

```{r}
read_excel(myExcelPath, range = "C1:E4")
```

```{r}
read_excel(myExcelPath, range = cell_cols("B:D"))
```

```{r}
read_excel(myExcelPath, range = "mtcars!B1:D5")
```

---

## Write to an Excel file

tidyverse doesn't have a package for writing to Excel files. We will use `write.xlsx()` from **openxlsx**. This package needs to be installed and loaded. Follow the instructions on the setup notebook for installation:

```{r}
library(openxlsx)  # for writing to Excel files
```

Let's write some of the data frames we have worked with into an Excel file:

```{r}
l <- list(mpg=mpg, diamonds=diamonds, iris=iris)
openxlsx::write.xlsx(l, file = "myNewDatasets.xlsx")
```

Reminder: `openxlsx::` is optional, just a good practice for showing that unlike all of the other packages used in this notebook `write.xlsx()` is not from tidyverse.

---

### Other types of data
* **haven** reads SPSS, Stata, and SAS files
* **DBI**, along with a database specific backend (e.g. RMySQL, RSQLite, RPostgreSQL etc) allows you to run SQL queries against a database and return a data frame
* For hierarchical data: use **jsonlite** for json, and **xml2** for XML
* **bigrquery** package allows read/write from/to BigQuery tables.

